---
title: "rdwd use cases"
author: "Berry Boessenkool, <berry-b@gmx.de>"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{rdwd use cases}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

[Vignette Rmd source code](https://raw.githubusercontent.com/brry/rdwd/master/localtests/CreateVignettes/cases.Rmd)
(Not on CRAN to reduce load on DWD server through daily new builds and checks of the vignette)

[main rdwd vignette](rdwd.html)

[Interactive map vignette](mapDWD.html)


## Get all hourly rainfall data 2014:2016

### Step 1: get the URLS of data to be downloaded

```{r hourlyrain_data_selection, warning=FALSE}
library(rdwd)
links <- selectDWD(res="daily", var="more_precip", per="hist")
length(links) # 5583 stations - would take very long to download

# select only the relevant files:
data("metaIndex")
myIndex <- metaIndex[
  metaIndex$von_datum < 20140101 &
  metaIndex$bis_datum > 20161231 & metaIndex$hasfile   ,  ]
data("fileIndex")    
links <- fileIndex[
  suppressWarnings(as.numeric(fileIndex$id)) %in% myIndex$Stations_id &
  fileIndex$res=="daily" &
  fileIndex$var=="more_precip" &
  fileIndex$per=="historical"         , "path" ]  
links <- paste0("ftp://ftp-cdc.dwd.de/pub/CDC/observations_germany/climate/", links)

length(links) # 2001 elements - much better
```


### Step 2: download the data

In real life, use a permanent folder instead of `tempdir()`.
If some downloads fail (mostly because you'll get kicked off the FTP server),
you can just run the same code again and only the missing files will be downloaded.

If you really want to download 2k historical (large!) datasets, 
you definitely want to set `sleep` to a much higher value.

For speed, we'll only work with the first 3 urls.

```{r hourlyrain_data_download, message=FALSE}
localfiles <- dataDWD(links[1:3], dir=tempdir(), sleep=0.2, read=FALSE)
```


### Step 3: read the data

2k large datasets probably is way too much for memory, so we'll use a custom reading function.
It will only select the relevant time section and rainfall column.
The latter will be named with the id extracted from the filename.

```{r hourlyrain_reading_function}
readVars(localfiles[1])[,-3] # we want the RS column

read2014_2016 <- function(file, fread=TRUE, ...)
{
 out <- readDWD(file, fread=fread, ...)
 out <- out[out$MESS_DATUM > as.POSIXct(as.Date("2014-01-01")) & 
            out$MESS_DATUM < as.POSIXct(as.Date("2016-12-31"))    , ]
 out <- out[ , c("MESS_DATUM", "RS")]
 out$MESS_DATUM <- as.Date(out$MESS_DATUM) # might save some memory space...
 # Station id as column name:
 idstringloc <- unlist(gregexpr(pattern="tageswerte_RR_", file))
 idstring <- substring(file, idstringloc+14, idstringloc+18)
 colnames(out) <- c("date",  idstring)
 return(out)
}
str(read2014_2016(localfiles[1])) # test looks good
```

Now let's apply this to all our files and merge the result.

```{r hourlyrain_data_reading}
library(pbapply) # progress bar for lapply loop

rain_list <- pblapply(localfiles, read2014_2016)
rain_df <- Reduce(function(...) merge(..., all=T), rain_list)
str(rain_df) # looks nice!
summary(rain_df) # 9 NAs in station 00006
```


[TOC](#top)

### step 4: visual data checks

```{r hourlyrain_vis}
plot(rain_df$date, rain_df[,2], type="n", ylim=range(rain_df[,-1], na.rm=T), 
     las=1, xaxt="n", xlab="Date", ylab="Daily rainfall sum  [mm]")
berryFunctions::monthAxis()
for(i in 2:ncol(rain_df)) lines(rain_df$date, rain_df[,i], col=sample(colours(), size=1))

plot(rain_df[,2:4]) # correlation plot only works for a few columns!
```

Let's see the locations of our stations in an interactive map.

```{r hourlyrain_map_interactive, warning=FALSE}
if(requireNamespace("leaflet", quietly=TRUE)){
data(geoIndex)  ;  library(leaflet) 
mygeoIndex <- geoIndex[geoIndex$id %in% as.numeric(colnames(rain_df)[-1]),]

leaflet(data=mygeoIndex) %>% addTiles() %>%
        addCircleMarkers(~lon, ~lat, popup=~display, stroke=T)
}
```    

For a static map with scaleBar, OSMscale works nicely but currently still has a Java dependency, see
<https://github.com/brry/OSMscale#installation>

```{r hourlyrain_map_static, message=FALSE}
if(requireNamespace("OSMscale")){
library(OSMscale)
pointsMap("lat", "lon", mygeoIndex, fx=2, fy=1, pargs=list(lwd=3), 
                    col="blue", zoom=5)
}
```    


[TOC](#top)

Any feedback on this package (or this vignette) is very welcome via 
[github](https://github.com/brry/rdwd) or <berry-b@gmx.de>!
